\LaTeX\ Author Guidelines for ICCV Proceedings

### 2021-03-02 chapters/0_abstract.tex
Shuffle stuff around; update intro

% !TEX root = ../agglo_clust_review.tex
%%%%%%%%% ABSTRACT
\begin{abstract}
\TODO{Update} We propose a new theoretical framework that generalizes algorithms for hierarchical agglomerative clustering to weighted graphs with both attractive and repulsive interactions between the nodes. This framework defines GASP, a Generalized Algorithm for Signed graph Partitioning, and allows us to explore many combinations of different linkage criteria and cannot-link constraints. We prove the equivalence of existing clustering methods to some of those combinations, and introduce new algorithms for combinations which have not been studied before. 
On stochastic block model problems, GASP compares favorably to spectral clustering. More importantly, we conduct a systematic comparison of various instantiations of GASP in image instance segmentation problems, in terms of accuracy but also efficiency and robustness to noise. 
We find that one of the new algorithms proposed in our framework outperforms all previously known agglomerative methods for signed graphs, both on the competitive CREMI 2016 EM segmentation benchmark and on the CityScapes dataset.
% \keywords{Partitioning of Signed Graphs, Generalized Framework, Agglomerative Clustering, Instance Segmentation}
\end{abstract}

>>>>

% !TEX root = ../agglo_clust_review.tex
%%%%%%%%% ABSTRACT
\begin{abstract}
We propose a new simple theoretical framework that generalizes algorithms for hierarchical agglomerative clustering to weighted graphs with both attractive and repulsive interactions between the nodes. This framework defines GASP, a Generalized Algorithm for Signed graph Partitioning, and allows us to explore many combinations of different linkage criteria and cannot-link constraints. 
We prove the equivalence of existing clustering methods to some of those combinations, and introduce new algorithms for combinations which have not been studied before. 
We also study theoretical properties of these combinations and prove which one of them define ultrametrics on the graph.
% On stochastic block model problems, GASP compares favorably to spectral clustering. 
We conduct a systematic comparison of various instantiations of GASP on a large variety of both synthetic and existing signed clustering problems, in terms of accuracy but also efficiency and robustness to noise. 
More importantly, we find that some of the algorithms proposed in our framework, when combined to the predictions from a CNN model, result in a simple instance segmentation pipeline that achieves state-of-the-art performances on the competitive CREMI 2016 EM segmentation benchmark.
% \keywords{Partitioning of Signed Graphs, Generalized Framework, Agglomerative Clustering, Instance Segmentation}
\end{abstract}

### 2021-03-02 chapters/1_intro.tex
Shuffle stuff around; update intro

In this paper, we propose a new theoretical framework that generalizes over agglomerative algorithms for signed graphs by linking them to hierarchical agglomerative clustering on positive-weighted graphs \cite{lance1967general}. This framework defines an underlying basic algorithm and allows us to explore its combinations with different linkage criteria and \emph{cannot-link constraints}. 
We then formally prove that some of the combinations correspond to existing clustering algorithms and introduce new algorithms for combinations which have not been explored before.

We evaluate and compare these algorithms on \emph{instance segmentation} -- a computer vision task of assigning each pixel of an image to an object instance. 
We use a CNN to predict the edge weights of a graph such that each node represents a pixel of the image, similarly to \cite{liu2018affinity,lee2017superhuman,wolf2018mutex}, and provide these weights as input to the algorithms in our framework (see Fig.~\ref{fig:intro_figure}).

@@

        \end{tabular}
    \end{subtable} 
    \caption{Clustering algorithms included in the proposed \algname{} framework, given a linkage criterion, a type of graph (signed or unsigned) and the optional use of cannot-link constraints. New algorithms are highlighted in yellow. Algorithms typeset in bold font define an ultrametric on the graph (see Eq.~\ref{eq:UM_def}). Algorithms included in the green box, are weight-shift invariant (see Prop.~\ref{prop:weight_shift_invariant}). 
    We denote as $E_{ij}=(S_i \times S_{j}) \cap E$ the set of edges connecting two cluster $S_i, S_j$. } 
    \label{tab:linkage-criteria}
\end{table*}

With our comparison experiments, performed both on 2D urban scenes from the CityScapes dataset and 3D electron microscopy image volumes of neurons, we benchmark all algorithms in our framework, focusing on their efficiency, robustness and tendency to over- or under-cluster.
We show that one of the new algorithms derived from our framework, based on an average linkage criterion, outperforms all previously known agglomeration methods expressed in the framework and that
it achieves competitive performance on CityScapes and the challenging CREMI 2016 segmentation benchmark.

In Sec.~\ref{sec:spectral_clust}, we also show how \algname{} outperforms spectral clustering methods on the task of neuron segmentation and how on synthetic graphs it achieves similar scores to a recently proposed spectral method for signed graphs. \TODO{Define HC and AC abbrv}

>>>>

@@

        \end{tabular}
    \end{subtable} 
    \caption{Clustering algorithms included in the proposed \algname{} framework, given a linkage criterion, a type of graph (signed or unsigned) and the optional use of cannot-link constraints. New algorithms are highlighted in yellow (hierarchical clustering with constraints, HCC). Algorithms typeset in bold font define an ultrametric on the graph (see Eq.~\ref{eq:UM_def}). Algorithms included in the green box, are weight-shift invariant (see Prop.~\ref{prop:weight_shift_invariant}). 
    We denote as $E_{ij}=(S_i \times S_{j}) \cap E$ the set of edges connecting two clusters $S_i, S_j$. } 
    \label{tab:linkage-criteria}
\end{table*}

In this paper, we propose a simple theoretical framework that generalizes over agglomerative algorithms for signed graphs by linking them to hierarchical clustering (HC) on positive-weighted graphs \cite{lance1967general}. This framework defines an underlying basic algorithm and allows us to explore its combinations with different linkage criteria and \emph{cannot-link constraints} (see Table~\ref{tab:linkage-criteria}). 
We then formally prove that some of the combinations correspond to existing clustering algorithms, and introduce new algorithms for combinations which have not been explored before. By analyzing their theoretical properties, we also show which one of them define ultrametrics on the graph.

We evaluate the algorithms on a large variety of existing and synthetic signed graph clustering problems. 
Our experiments show that the choice of a linkage criteria strongly influence the way how clusters are grown by the agglomerative algorithms, making some linkage methods more suited for certain types of clustering problems.
% We focus our comparison on three of the best-performing algorithms included in our framework, which are based on average, sum, and absolute maximum linkage criteria (see Table~\ref{tab:linkage-criteria}). We highlight their empirical properties by showing how clusters are grown differently depending on the algorithm.
Moreover, we evaluate all algorithms on \emph{instance segmentation} -- a computer vision task consisting in assigning each pixel of an image to an object instance -- by predicting graph edge weights with a CNN (see Fig.~\ref{fig:intro_figure}).
We benchmark the clustering algorithms by focusing on their efficiency, robustness and tendency to over- or under-cluster. 
First, we observe that the tested agglomerative algorithms strongly outperform recently proposed spectral clustering methods. Then,  we also show that average-linkage based agglomerative algorithms achieve state of the art results on the competitive CREMI 2016 challenge for neuron segmentation of 3D electron microscopy image volumes of brain tissue.

% In Sec.~\ref{sec:spectral_clust}, we also show how \algname{} outperforms spectral clustering methods on the task of neuron segmentation and how on synthetic graphs it achieves similar scores to a recently proposed spectral method for signed graphs. 
% \TODO{Define HC abbrv}

### 2021-03-02 chapters/2_rel_work.tex
Shuffle stuff around; update intro

\section{Related work} \label{sec:related_work}
\textbf{Proposal-based methods} have been highly successful in instance segmentation competitions like MS COCO \cite{lin2014microsoft}, Pascal VOC2012 \cite{everingham2010pascal} and CityScapes \cite{cordts2016cityscapes}. They decompose the instance segmentation task into two steps that consists in generating object proposals and assigning to each bounding box a class and a binary segmentation mask \cite{he2017mask,porzi2019seamless,liu2018path,yang2012layered,li2017fully,ladicky2010and,hariharan2014simultaneous,chen2015multi,dai2016instance,liang2016reversible}. 
% They commonly rely on {Faster-RCNN}~\cite{ren2015faster} and can be trained end-to-end using non-maximum suppression. 
Other methods use instead recurrent models to sequentially generate instances one-by-one \cite{romera2016recurrent,ren2017end}.

\textbf{Proposal-free methods} adopt a bottom-up approach by directly grouping pixels into instances. Recently, there has been a growing interest for such  methods that do not involve object detection, since, in certain types of data, object instances cannot be approximated by bounding boxes. For example, the approach proposed in \cite{kirillov2017instancecut} uses a combinatorial framework for instance segmentation, 
% SGN \cite{liu2017sgn} sequentially group pixels into lines and then instances;
whereas a watershed transform is learned in \cite{bai2017deep} by also predicting its gradient direction. 
% whereas the template matching \cite{uhrig2016pixel} deploys scene depth information.
Others use metric learning to predict high-dimensional associative pixel embeddings that map pixels of the same instance close to each other, while mapping pixels belonging to different instances further apart \cite{lee2019learning,fathi2017semantic,newell2017associative,de2017semantic}. % kulikov2018instance
Final instances are then retrieved by applying a clustering algorithm, like in the end-to-end trainable mean-shift pipeline of \cite{kong2018recurrentPix}. 
Other recent successful methods simply let the model predict the relative coordinates of the instance center \cite{neven2019instance,cheng2019panopticdeeplab} or, given a point $(x,y)$ in the image, they train a model to generate the mask of the instance located at $(x,y)$ \cite{sofiiuk2019adaptis}.

\textbf{Edge detection} also experienced recent progress thanks to deep learning, both on natural images \cite{Gao_2019_ICCV,liu2018affinity,xie2015holistically,kokkinos2015pushing} and biological data \cite{lee2017superhuman,schmidt2018cell,meirovitch2016multi,ciresan2012deep}. In neuron segmentation for connectomics, a field of neuroscience we also address in our experiments, boundaries are converted to final instances with subsequent postprocessing and superpixel-merging:
some use loopy graphs \cite{kaynig2015large,krasowski2015improving} or trees \cite{meirovitch2016multi,liu2016sshmt,liu2014modular,funke2015learning,uzunbas2016efficient} to represent the region merging hierarchy; the lifted multicut \cite{beier2017multicut} formulates the problem in a combinatorial framework, whereas 
flood-filling networks \cite{januszewski2018high} and MaskExtend \cite{meirovitch2016multi} use a CNN to iteratively grow one region/neuron at the time; recently, the work of \cite{meirovitch2019cross} made the process more efficient by employing a combinatorial encoding of the segmentation.
A structured learning approach was also proposed in \cite{funke2018large,turaga2009maximin}.

@@

This work reformulates the clustering algorithms of \cite{levinkov2017comparative,wolf2018mutex,keuper2015efficient} in a generalized framework and adopt ideas from the proposal-free methods \cite{liu2018affinity,wolf2018mutex,lee2017superhuman} to predict long-range relationships between pixels.

>>>>

\section{Related work} \label{sec:related_work}
\textbf{Proposal-based methods} have been highly successful in instance segmentation competitions like MS COCO \cite{lin2014microsoft}and CityScapes \cite{cordts2016cityscapes}. They decompose the instance segmentation task into two steps that consists in generating object proposals and assigning to each bounding box a class and a binary segmentation mask \cite{he2017mask,porzi2019seamless,liu2018path,yang2012layered,li2017fully,ladicky2010and,hariharan2014simultaneous,chen2015multi,dai2016instance,liang2016reversible}. 
% They commonly rely on {Faster-RCNN}~\cite{ren2015faster} and can be trained end-to-end using non-maximum suppression. 
% Other methods use instead recurrent models to sequentially generate instances one-by-one \cite{romera2016recurrent,ren2017end}.

\textbf{Proposal-free methods} adopt a bottom-up approach by directly grouping pixels into instances. Recently, there has been a growing interest for such  methods that do not involve object detection, since, in certain types of data, object instances cannot be approximated by bounding boxes \cite{kirillov2017instancecut,bai2017deep}. 
% For example, the approach proposed in \cite{kirillov2017instancecut} uses a combinatorial framework for instance segmentation, 
% SGN \cite{liu2017sgn} sequentially group pixels into lines and then instances;
% whereas a watershed transform is learned in \cite{bai2017deep} by also predicting its gradient direction. 
% whereas the template matching \cite{uhrig2016pixel} deploys scene depth information.
Some use metric learning to predict high-dimensional associative pixel embeddings that map pixels of the same instance close to each other \cite{lee2019learning,fathi2017semantic,newell2017associative,de2017semantic}
% , while mapping pixels belonging to different instances further apart \cite{lee2019learning,fathi2017semantic,newell2017associative,de2017semantic}. % kulikov2018instance
and then retrieve final instances by applying a clustering algorithm \cite{kong2018recurrentPix}.
% , like in the end-to-end trainable mean-shift pipeline of \cite{kong2018recurrentPix}. 
Other recent successful methods simply let the model predict the relative coordinates of the instance center \cite{neven2019instance,cheng2019panopticdeeplab} or, given a point $(x,y)$ in the image, they train a model to generate the mask of the instance located at $(x,y)$ \cite{sofiiuk2019adaptis}.

\textbf{Edge detection} also experienced recent progress thanks to deep learning, both on natural images \cite{Gao_2019_ICCV,liu2018affinity,xie2015holistically,kokkinos2015pushing} and biological data \cite{lee2017superhuman,schmidt2018cell,meirovitch2016multi,ciresan2012deep}. In neuron segmentation for connectomics, a field of neuroscience we also address in our experiments, boundaries are converted to final instances with subsequent postprocessing and superpixel-merging:
some use a combinatorial framework \cite{beier2017multicut}, others use loopy graphs \cite{kaynig2015large,krasowski2015improving} or trees \cite{meirovitch2016multi,liu2016sshmt,liu2014modular,funke2015learning,uzunbas2016efficient} to represent the region merging hierarchy. Flood-filling networks \cite{januszewski2018high} and MaskExtend \cite{meirovitch2016multi} used a CNN to iteratively grow one region/neuron at the time.
% recently, the work of \cite{meirovitch2019cross} made the process more efficient by employing a combinatorial encoding of the segmentation.
A structured learning approach was also proposed in \cite{funke2018large,turaga2009maximin}.

@@

This work reformulates the clustering algorithms of \cite{levinkov2017comparative,wolf2018mutex,keuper2015efficient} in a generalized framework and adopt ideas from the proposal-free instance segmentation methods \cite{liu2018affinity,wolf2018mutex,lee2017superhuman} to predict long-range affinities between pixels.

### 2021-03-02 chapters/3_algorithm.tex
Small changes
Shuffle stuff around; update intro

\subsection{New and existing algorithms included in \algname{}} \label{sec:alg_update_rules}

@@

\paragraph{Average, Single, and Complete Linkage} -- These three linkage criteria have been throughly studied on positive-weighted graphs. In \cite{chehreghani2020hierarchical}, the authors proved that applying these three linkage methods to a signed graph is equivalent to applying them to the positive-weighted graph obtained by shifting all edge weights by a constant.  Here, we prove which of the algorithms included in our generalized framework have this invariance-property: (proof in Appendix)
\begin{prop} \label{prop:weight_shift_invariant}
We call an agglomerative algorithm ``weight-shift invariant'' if the structure of the merging tree $T$ returned by such algorithm is invariant w.r.t. a shift of all edge weights $w_e$  by a constant $\alpha\in \mathbb{R}$. Among the versions of \algname{} studied in this paper, only those using Average, Single, and Complete linkage criteria without cannotLinkConstraints are weight-shift-invariant (see algorithms typeset in bold font in Table~\ref{tab:linkage-criteria}).
\end{prop}
\noindent To our knowledge, Avg, Single, and Complete linkage methods have never been combined with cannot-link constraints on signed graphs, so the algorithms that in Table~\ref{tab:linkage-criteria} are named \emph{HCC-Avg}, \emph{HCC-Single}, and \emph{HCC-Complete} are new.

@@

\paragraph{Algorithms defining an ultrametric} -- The connection between agglomerative algorithms and ultrametrics\footnote{A metric space $(X,d)$ is an \emph{ultrametric} if for every $x,y,z \in X$, $d(x,y)\leq \max \{d(x,z), d(y,z)\}$.} is well known \cite{johnson1967hierarchical,milligan1979ultrametric}. Normally, ultrametrics are associated to strictly positive \emph{similarity} or \emph{dissimilarity} measures on a graph. Here, we generalize this concept to signed graphs. To do so, we first map the signed interaction $\treeHeight{}$ defined in Eq.~\ref{eq:def_dendr_interact} to positive ``distances'' $d_{T}:V \times V \rightarrow \mathbb{R}^{+}$:
\begin{align}\label{eq:UM_def}
d_{T}(u,v)\equiv &\begin{cases}
0 & \text{if}\,\, u=v\\
M-\treeHeight(u,v) & \text{if}\,\, u\neq v
\end{cases}\quad \forall u,v\in V\\
\mathrm{where}& \quad  M \equiv  \epsilon + \max_{u',v'\in V,\,u'\neq v'}\treeHeight(u',v')
\end{align}
and where $\epsilon > 0$. By definition, we have that $d_{T}(u,v)=d_{T}(v,u)$ and $d_{T}(u,u)= 0$. We then prove the following two propositions (proofs are left in Appendix):
\begin{prop}
Consider a graph $\mathcal{G}(V,E,w_e)$, a linkage criteria $\interact{}$, and an AC algorithm returning a binary rooted tree $T$ with height $h_T$. The quantity $(V, d_{T})$ defined in Eq.~\ref{eq:UM_def} is an ultrametric iff, $\forall u,v,t \in V$:
\begin{equation}
\interact{}_{T}(u,v) > \interact{}_{T}(u,t) \Leftrightarrow h_T(u \vee v) < h_T[u \vee t] 
\end{equation}
In words, $(V, d_{T})$ is an ultrametric iff the interaction $\interact{}_{T}(u, v)$ associated to the least common ancestor ${(u\vee v)\in T}$ is both lower than any other $(u\vee v)$-descendant node and higher than any other $(u\vee v)$-ancestor node in the tree $T$.
\end{prop}
\begin{prop}
% Consider the binary rooted tree $T^*$ returned by the \TODO{all variants in Table 1} \algname{} Algorithm~\ref{main_alg}, given a choice of linkage criteria $\interact{}$ and the use (or not) of CannotLinkConstraints. 
Among the algorithms listed in Table \ref{tab:linkage-criteria} and included in the \algname{} framework, only Mutex Watershed and hierarchical clustering with Average (HC-Avg), Single (HC-Single) and Complete linkage (HC-Complete) define an ultrametric given by the quantity $(V, d_{T^*})$ defined in Eq.~\ref{eq:UM_def}, where $T^*$ is the tree returned by Algorithm~\ref{main_alg}.
\end{prop}

>>>>

\subsection{\algname{}: New and existing algorithms} \label{sec:alg_update_rules}

@@

\paragraph{Average, Single, and Complete Linkage} -- These three linkage criteria have been throughly studied on positive-weighted graphs, but never on signed graphs. In the recent related work \cite{chehreghani2020hierarchical}, the authors prove that applying these three linkage methods to a signed graph is equivalent to applying them to the positive-weighted graph obtained by shifting all edge weights by a constant.  Here, we prove which of the algorithms included in our generalized framework have this invariance-property: (proof in Appendix)
\begin{prop} \label{prop:weight_shift_invariant}
We call an agglomerative algorithm ``weight-shift invariant'' if the structure of the merging tree $T$ returned by such algorithm is invariant w.r.t. a shift of all edge weights $w_e$  by a constant $\alpha\in \mathbb{R}$. Among the versions of \algname{} studied in this paper, only those using Average, Single, and Complete linkage criteria without cannotLinkConstraints are weight-shift-invariant (see algorithms typeset in bold font in Table~\ref{tab:linkage-criteria}).
\end{prop}
\noindent To our knowledge, Avg, Single, and Complete linkage methods have never been combined with cannot-link constraints on signed graphs, so the algorithms that in Table~\ref{tab:linkage-criteria} are named \emph{HCC-Avg}, \emph{HCC-Single}, and \emph{HCC-Complete} are new.

@@

\paragraph{Algorithms defining an ultrametric} -- The connection between agglomerative algorithms and ultrametrics\footnote{A metric space $(X,d)$ is an \emph{ultrametric} if for every $x,y,z \in X$, $d(x,y)\leq \max \{d(x,z), d(y,z)\}$.} is well known \cite{johnson1967hierarchical,milligan1979ultrametric}. Usually, ultrametrics are associated to strictly positive \emph{similarity} or \emph{dissimilarity} measures on a graph. Here, we generalize this concept to signed graphs. To do so, we first map the signed interaction $\treeHeight{}$ defined in Eq.~\ref{eq:def_dendr_interact} to positive ``distances'' $d_{T}:V \times V \rightarrow \mathbb{R}^{+}$:
\begin{align}\label{eq:UM_def}
d_{T}(u,v)\equiv &\begin{cases}
0 & \text{if}\,\, u=v\\
M-\treeHeight(u,v) & \text{if}\,\, u\neq v
\end{cases}\quad \forall u,v\in V\\
\mathrm{where}& \quad  M \equiv  \epsilon + \max_{u',v'\in V,\,u'\neq v'}\treeHeight(u',v')
\end{align}
and where $\epsilon > 0$. By definition, we have that $d_{T}(u,v)=d_{T}(v,u)$ and $d_{T}(u,u)= 0$. We then prove the following two propositions (proofs are left in Appendix):
\begin{prop}
Consider a graph $\mathcal{G}(V,E,w_e)$, a linkage criteria $\interact{}$, and an AC algorithm returning a binary rooted tree $T$ with height $h_T$. The quantity $(V, d_{T})$ defined in Eq.~\ref{eq:UM_def} is an ultrametric iff, $\forall u,v,t \in V$:
\begin{equation}
\interact{}_{T}(u,v) > \interact{}_{T}(u,t) \Leftrightarrow h_T(u \vee v) < h_T[u \vee t] 
\end{equation}
In words, $(V, d_{T})$ is an ultrametric iff the interaction $\interact{}_{T}(u, v)$ associated to the least common ancestor ${(u\vee v)\in T}$ is both lower than any other $(u\vee v)$-descendant node and higher than any other $(u\vee v)$-ancestor node in the tree $T$.
\end{prop}
\begin{prop}
% Consider the binary rooted tree $T^*$ returned by the \TODO{all variants in Table 1} \algname{} Algorithm~\ref{main_alg}, given a choice of linkage criteria $\interact{}$ and the use (or not) of CannotLinkConstraints. 
Among the algorithms listed in Table \ref{tab:linkage-criteria} and included in the \algname{} framework, only Mutex Watershed and hierarchical clustering with Average (HC-Avg), Single (HC-Single) and Complete linkage (HC-Complete) define an ultrametric given by the quantity $(V, d_{T^*})$ defined in Eq.~\ref{eq:UM_def}, where $T^*$ is the tree returned by Algorithm~\ref{main_alg}.
\end{prop}

### 2021-03-02 chapters/5_experiments_compare_rules.tex
Shuffle stuff around; update intro

\begin{figure}[t]
\centering
\includegraphics[width=0.47\textwidth,trim=0.25in 0.in 0.5in 0.in,clip]{./figs/SSBM_experiments.pdf}
        \caption{Scores on SSBM graphs}
\end{figure}
\begin{figure}[t]
\centering
\includegraphics[width=0.47\textwidth,trim=0.53in 0.1in 0.65in 0.45in,clip]{./figs/noise_plots/under_segment_plots_1.pdf}
        \caption{Scores on \emph{CREMI-gridGraph} perturbed with noise}
\end{figure}

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=0.47\textwidth,trim=0.53in 0.1in 0.65in 0.45in,clip]{./figs/noise_plots/under_segment_plots_1.pdf}
        \caption{
ARAND errors (lower is better) on \emph{CREMI-gridGraph} clustering problems perturbed with structured noise. Average-linkage algorithms proved to be the most robust.
We report median, 25th, and the 75th percentile values over \TODO{30} experiments.
% \algname{} performances compared to spectral methods on synthetic graphs. The spectral methods were given the true number of clusters as input, in contrast to \algname{}. \TODO{Define paramters SSB; explain percentile stuff (for both noise experiments), user ARAND-error}
}\label{fig:scores_structured_noise}
\end{figure}
